The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-27 manuscript.tex
update tex file

% !TeX root = RJwrapper.tex
\title{The Rockerverse: Images, Packages, and Applications for Containerization
with R}
\author{by Daniel Nüst, Dom Bennett, Robrecht Cannoodt, Dav Clark, Gergely Daroczi, Dirk Eddelbuettel, Mark Edmondson, Colin Fay, Ellis Hughes, Sean Lopp, Ben Marwick, Heather Nolis, Jacqueline Nolis, Hong Ooi, Karthik Ram, Noam Ross, Lori Shepherd, Nitesh Turaga, Craig Willis, Nan Xiao, Charlotte Van Petegem}

@@

\textbf{Whole~Tale} (\url{https://wholetale.org}) is a National Science
Foundation (NSF) funded project developing a scalable, open-source,
multi-user platform for reproducible research
\citep{brinckman2019, chard2019a}. A central goal of the platform is to
enable researchers to easily create and publish executable research
objects\textbackslash{}footnote\{In Whole\textasciitilde{}Tale a
\emph{tale} is a research object that contains metadata, data (by copy
or reference), code, narrative, documentation, provenance, and
information about the computational environment to support computational
reproducibility\} associated with published research \citep{chard2019b}.
Using Whole~Tale, researchers can create and publish Rocker-based
reproducible research objects to a growing number of repositories
including DataONE member nodes, Zenodo and soon Dataverse. Additionally,
Whole~Tale supports automatic data citation and is working on
capabilities for image preservation and provenance capture to improve
the transparency of published computational research artifacts
\citep{mecum2018, mcphillips2019}. For R users, Whole~Tale attempts to
combine the strengths of the Rocker Project's curated Docker images with
the Jupyter Project's
\texttt{repo2docker}\footnote{See \href{https://repo2docker.readthedocs.io/}{https://repo2docker.readthedocs.io/}}
component to simplify the customization of R-based environments for
researchers with limited experience with either Docker or Git. Multiple
options have been discussed to allow users to change the base image used
in \texttt{repo2docker} from the default Ubuntu LTS (long-term support)
required to support the Rocker Project images. Whole~Tale implemented a
custom
\texttt{RockerBuildPack}\footnote{See \href{https://github.com/whole-tale/repo2docker\_wholetale}{https://github.com/whole-tale/repo2docker\_wholetale}}
to support customization of the \texttt{rocker/geospatial} image through
\texttt{repo2docker}
composability\footnote{Composability refers to the ability to combine multiple package managers -- such as R, `pip`, and `conda`}.
This works in part because Rocker images are based on a Debian
distribution, so the instructions created by \texttt{repo2docker} for
Ubuntu work because of compatible shell and package manager.

@@

\textbf{\CRANpkg{stevedore}}
(\url{https://cran.r-project.org/package=stevedore}) is currently the
most powerful Docker client in R. It interfaces with the Docker daemon
over the Docker HTTP
API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}}
via a Unix socket on Linux or MacOS, over a named pipe on Windows, or
over an HTTP/TCP connection. The package is the only one not using
system calls to the \texttt{docker} CLI for managing images and
containers and easily exposes connections to remote Docker daemons,
which has to be configured on the Docker level otherwise. Using the API
gives access to more information and is system independent and likely
more reliable than parsing command line output. \CRANpkg{stevedore}'s
own interface is automatically generated based on the OpenAPI
specification of the Docker daemon, but still similar to the Docker CLI.
The interface is similar to R6 objects, in that a
\texttt{stevedore\_object} has a number of functions attached to it that
can be called, and multiple specific versions of the Docker API can be
supported thanks to the automatic
generation\footnote{See \href{https://github.com/richfitz/stevedore/tree/master/inst/spec}{https://github.com/richfitz/stevedore/tree/master/inst/spec} and \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

\textbf{\CRANpkg{AzureContainers}} is an interface to a number of
container-related services in Microsoft's
\href{https://azure.microsoft.com/}{Azure Cloud}, see
Section~\nameref{azurecontainers}. While it is mainly intended for
working with Azure, as a convenience feature it includes lightweight,
cross-platform shells to Docker and Kubernetes (tools \texttt{kubectl}
and \texttt{helm}). These can be used to create and manage arbitrary
Docker images and containers, as well as Kubernetes clusters on any
platform or cloud service.

@@

CARD.com implemented a web application for the optimization of the
acquisition flow and the real-time analysis of debit card transactions.
The software used \CRANpkg{Rserve} and rApache and was deployed in
Docker containers. The R session behind \CRANpkg{Rserve} acted as a
read-only in-memory database, which was extremely fast and scalabale,
for the many concurrent rApache processes responding to the live-scoring
requests of various divisions of the company. Similarly dockerized R
scripts were responsible for the ETL processes and even the
client-facing email, text message and push notification alerts sent in
real-time based on card transactions. The related Docker images were
made available at \url{https://github.com/cardcorp/card-rocker}. The
images extend \texttt{rocker/r-base} and additionally entailed an SSH
client and a workaround for being able to mount SSH keys from the host,
Pandoc, the Amazon Web Services (AWS) SDK, and Java, which is required
by the AWS SDK. The AWS SDK allowed to run R consumers reading from
real-time data processing streams of
\href{https://aws.amazon.com/kinesis/}{AWS
Kinesis}\textbackslash{}footnote\{See useR!2017 talk \emph{Stream
processing with R in AWS} at
\textbackslash{}href\{\url{https://static.sched.com/hosted_files/user2017/2f/AWR\%20Kinesis\%20at\%20useR\%202017.pdf}\}\{\url{https://static.sched.com/hosted_files/user2017/2f/AWR}
Kinesis at use 02017.pdf\}.\}. The applications were deployed on Amazon
Elastic Container Service (\href{https://aws.amazon.com/ecs/}{ECS}). The
main learnings from using R in Docker was the imporance of not only
pinning the R package versions via MRAN, but also moving away from
Debian testing to a distribution with long-term support. For the use
case at hand, this switch served the priority to have more control over
upstream updates, and to minimize the risk of breaking the automated
builds of the Docker images and production jobs.

@@

\hypertarget{continuous-integration-continuous-delivery-and-deployment}{%
\subsection{Continuous integration, continuous delivery, and
deployment}\label{continuous-integration-continuous-delivery-and-deployment}}

\label{cicd} \label{deployment}

The cloud is the natural environment of containers, and subsequently
containers are the go-to mechanism to deploy R server applications. More
and more continuous integration (CI) and continuous delivery (CD)
services also use containers, opening up new options for use. The
controlled nature of containers, i.e., the possibility to define the
software environment very well, even on remote machines, is crucial
here, for example to match test or build environments with production
environments.

As a first example, \CRANpkg{Shiny} is the most popular package for
creating interactive online dashboards with R and it enables users with
very diverse backgrounds to create stable and user friendly web
applications. The ShinyProxy (\url{https://www.shinyproxy.io/}) is an
open source tool to deploy Shiny apps in an enterprise context. They
feature single sign-on, but also in scientific use cases
\citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to
achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate
requirements and may itself run in a container for stability and
availability. The tool is build on ContainerProxy
(\url{https://www.containerproxy.io/}), which provides similar features
for executign long-running R jobs or interactive R sessions. The started
containers can run on a regular Docker host but also in clusters.

@@

\begin{itemize}
\tightlist
\item
  \href{https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml}{dynwrap}
  \citep{rcannood}

  \begin{itemize}
  \tightlist
  \item
    For this project, we use travis-ci to build rocker-derived
    containers, test them, and only push them to Docker~Hub (from
    travis-ci.org) if the integration tests succeed.
  \end{itemize}
\end{itemize}

For example, the
\textbf{\href{https://github.com/ThinkR-open/golem}{\texttt{golem}}}
package makes an heavy use of \texttt{dockerfiler} when it comes to
creating the \texttt{Dockerfile} for building production-grade Shiny
applications and deploying them.

\begin{itemize}
\tightlist
\item
  \texttt{RSelenium}
\item
  \href{https://cloudyr.github.io/googleComputeEngineR/}{\texttt{googleComputeEngineR}}
  (function \texttt{gce\_vm\_template()})

  \begin{itemize}
  \tightlist
  \item
    To enable quick deployments of key R services such as RStudio and
    Shiny onto cloud virtual machines (VMs), this package utilises
    Dockerfiles to move the labour of setting up those services from the
    user to a premade Docker image. For example, by specifying the
    template \texttt{template="rstudio"} in
    \texttt{gce\_vm\_template()/gce\_vm()} an up to date RStudio Service
    image is launched. Specifying \texttt{template="rstudio-gpu"} will
    launch an RStudio Server image with a GPU attached, etc.\\
  \end{itemize}
\item
  \href{https://github.com/sckott/analogsea}{\texttt{analogsea}}
  (digital ocean R client)
\item
  \href{https://www.rplumber.io/docs/hosting.html\#docker}{\texttt{plumber}}
\item
  \href{https://github.com/tmobile/r-tensorflow-api}{production
  deployment of neural networks} \citep[\citet{nolistic}]{jnolis}
\item
  Kubernetes - A popular platform for managing Docker containers is
  \href{https://kubernetes.io/}{Kubernetes}, which is used for a wide
  variety of containerized applications. It may be your organisation has
  a Kubernetes cluster already for other applications. Docker containers
  are used within Kubernetes clusters to hold native code, for which
  Kubernetes creates a framework around network connections and scaling
  of resources up and down. An introduction on using
  \href{https://code.markedmondson.me/r-on-kubernetes-serverless-shiny-r-apis-and-scheduled-scripts/}{Kubernetes
  with R is published at this blog post}.
\end{itemize}

\label{azurecontainers} \textbf{\CRANpkg{AzureContainers}} is an
umbrella package providing interfaces three commercial services of
Microsoft's Azure Cloud, namely
\href{https://azure.microsoft.com/en-us/services/container-instances/}{Container
Instances} for running individual containers,
\href{https://azure.microsoft.com/en-us/services/container-registry/}{Container
Registry} for private image distribution, and
\href{https://azure.microsoft.com/en-us/services/kubernetes-service/}{Kubernetes
Service} for orchestrated deployments. While a package like
\CRANpkg{plumber} provides the infrastructure for turning an R workflow
a service, for production purposes it is usually necessary to take into
account scalability, reliability and ease of management. A
\emph{cluster} of containers, orchestrated as a single deployment, e.g.,
with \href{https://kubernetes.io}{Kubernetes}, can mitigate limitations
on request volumes or a container occupied with computationally
intensive task. A cluster features load-balancing, autoscaling of
containers across numerous servers (in the cloud or on premise), and
restarting failed ones. AzureContainers provides an R-based interface to
these features for Azure. On the client side, it provides simple
wrappers to Docker and to Kubernetes' management tools \texttt{kubectl}
and \texttt{helm}. The package simplifies complex infrastructure
management to a number of R function calls, given an Azure account with
sufficient
credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.

The prevelance of Docker in industry naturally leads to usage of
containers with R in such settings as well, as customers already manage
platforms in Docker containers. These products often entail a high
amount of open source software in combination with proprietary layers
addign the relevant commercializable features. One such example is
RStudio's data science platform
\href{https://rstudio.com/products/team/}{RStudio Team}. It allows teams
of data scientists and their respective IT/DevOps groups to develop and
deploy code in R and Python around the RStudio Open Source Server inside
of Docker images, without requiring users to learn new tools or directly
interact with containers. The best practices for
\href{https://support.rstudio.com/hc/en-us/articles/360021594513-Running-RStudio-with-Docker-containers}{running
RStudio with Docker containers} as well as
\href{https://github.com/rstudio/rstudio-docker-products}{Docker images}
for RStudio's commerical products are openly available.

@@

\textbf{\pkg{holepunch}} (\url{https://github.com/karthik/holepunch}) is
an R package that was designed to make sharing work environments
accessible to novice R users based on \textbf{Binder}. The
\href{https://mybinder.readthedocs.io/en/latest/}{Binder project},
maintained by the team behind Jupyter, makes it possible for users to
create and share their computing environments with others
\citep{jupyter_binder_2018}. A \emph{BinderHub} allows anyone with
access to a web browser and an internet connection to launch a temporary
instance of these custom environments and execute any workflows
contained within. From a reproducibility standpoint, Binder makes it
exceedingly easy to compile a paper, visualize data, and run small
examples from papers or tutorials without the need for any local
installation. To set up Binder for a project, a user typically starts at
an instance of a BinderHub and passes the location of a repository with
a workspace, e.g., a hosted Git repository, or a data repository like
Zenodo. Binder's core internal tool is
\href{https://repo2docker.readthedocs.io/en/latest/config_files.html}{\texttt{repo2docker}}.
It deterministically builds a Docker image by parsing the contents of a
repository, e.g., project dependency configurations or simple
configuration files. In the most powerful case, \texttt{repo2docker}
builds a given \texttt{Dockerfile}. While this approach works well for
most run of the mill Python projects, it is not so seamless for R
projects. For any R projects that use the Tidyverse suite
\citep{wickham_welcome_2019}, the time and resources required to build
all dependencies from source can often time out before completion,
making it frustrating for the average R user. \pkg{holepunch} removes
some of these limitations by leveraging Rocker images that contain the
Tidyverse along special Jupyter dependencies, and only installs
additional packages from CRAN and Bioconductor that are not already part
of these images. It short cicuits the configuration file parsing in
\texttt{repo2docker} and starts with the Binder/Tidyverse base images,
which eliminates a large part of the build time and in most cases
results in a binder instance launching within a minute. \pkg{holepunch}
as a side effect also creates a \texttt{DESCRIPTION} file which then
turns any project into a research compendium
\citep{marwick_packaging_2018}. The \texttt{Dockerfile} included with
the project can also be used to launch a RStudio server locally, i.e.,
independent of Binder, which is especially useful when more or special
computational resources can be provided there. The local image usage
reduces the number of seperately managed environments and thereby
reduces work and increases portability and reproducibility.

@@

The portability of containerized environments becomes particularly
useful for improving expensive processing of data or shipping complex
processing pipelines.

First, it is possible to \textbf{offload complex processing to a
server}. \ldots{}

Second, when processes can be \textbf{run in parallel} for speeding up R
code.

For example,
\href{https://CRAN.R-project.org/package=googleComputeEngineR}{\texttt{googleComputeEngineR}}'s
\texttt{gce\_vm\_cluster()} function can create clusters of 2 or more
virtual machines, running multi-CPU architectures. Instead of running a
local R script with the local CPU and RAM restrictions, the same code
can be processed on all CPU threads of the cluster of machines in the
cloud, all running in a Docker container with the same R environments.
This is achieved through \texttt{googleComputeEngineR}'s integration
with the R parralisation library
\href{https://CRAN.R-project.org/package=future}{\texttt{library(future)}
by Henrik Bengtsson}. Local R computation can be thrown up to a
multi-CPU and VM environment to achieve parrell computation in a few
lines of R code in your local session. -
\href{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}{some
demonstrations are available here}. - \href{https://cloud.run}{Google
Cloud Run} - This is a CaaS (Containers as a Service) that lets you
launch a Docker container without worrying about underlying
infrastructure. This dispenses with the developer creating a cloud
server to run the Docker image on, by abstracting away those servers to
a more serverless configuration. Cloud Run lets you run your code on top
of a managed or your own Kubernetes cluster running Knative, and can
accept any Docker image. The service takes care of network ingress,
scaling machines up and down to zero, authentication and authorisation,
all features which are non-trivial for a developer to create on their
own. This can be used to scale up R code to millions of instances if
they need to, with little or no changes to existing code. An R
implementation is shown here at
\href{https://github.com/MarkEdmondson1234/cloudRunR}{cloudRunR} which
uses Cloud Run to create a scalable R plumber API.
\url{https://code.markedmondson.me/googleCloudRunner/index.html}

\href{https://cloud.google.com/cloud-build/}{Google Cloud Build} and the
Google Container Registry are a continuous intergation service
respectively image registry that offload building of images to the
cloud, while serving the needs of commercial environments that helps
move the workload of building the Docker images to an online service,
while Cloud Build can be set up to build the Dockerfiles on each GitHub
commit or release. This means you do not need to build the Docker images
locally, which can tie up resources since Docker images can be several
GBs and take a long time to compile. Google Cloud Build works alongside
to allow you to build private and public Docker images, which allows you
to build up your own dependency graphs for downstream applications. -
\url{https://code.markedmondson.me/googleCloudRunner/articles/cloudbuild.html}

\begin{itemize}
\tightlist
\item
  \CRANpkg{batchtools} \citep{Lang2017batchtools} can
  \href{https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html}{schedule
  jobs with Docker Swarm}
\item
  scalable deployments, e.g., start with numerous Shiny talks mentioning
  Rocker at useR!2017
\item
  \href{https://github.com/dynverse/dynmethods}{dynmethods}
  \citep{rcannood}: In order to evaluate ±50 computational methods which
  all used different environments (R, Python, C++, \ldots{}), we wrapped
  each of them in a docker container and can execute these methods from
  R. Again, all of these containers are being built on travis-ci, and
  will only be pushed to Docker~Hub if the integration test succeeds.
\item
  \CRANpkg{drake},
  \url{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}
\end{itemize}

>>>>

% !TeX root = RJwrapper.tex
\title{The Rockerverse: Images, Packages, and Applications for Containerization
with R}
\author{by Daniel Nüst, Dirk Eddelbuettel, Dom Bennett, Robrecht Cannoodt, Dav Clark, Gergely Daroczi, Mark Edmondson, Colin Fay, Ellis Hughes, Sean Lopp, Ben Marwick, Heather Nolis, Jacqueline Nolis, Hong Ooi, Karthik Ram, Noam Ross, Lori Shepherd, Nitesh Turaga, Craig Willis, Nan Xiao, Charlotte Van Petegem}

@@

@@

\textbf{\CRANpkg{stevedore}}
(\url{https://cran.r-project.org/package=stevedore}) is currently the
most powerful Docker client in R. It interfaces with the Docker daemon
over the Docker HTTP
API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}}
via a Unix socket on Linux or MacOS, over a named pipe on Windows, or
over an HTTP/TCP connection. The package is the only one not using
system calls to the \texttt{docker} CLI for managing images and
containers and easily exposes connections to remote Docker daemons,
which has to be configured on the Docker level otherwise. Using the API
gives access to more information and is system independent and likely
more reliable than parsing command line output. \CRANpkg{stevedore}'s
own interface is automatically generated based on the OpenAPI
specification of the Docker daemon, but still similar to the Docker CLI.
The interface is similar to R6 objects, in that a
\texttt{stevedore\_object} has a number of functions attached to it that
can be called, and multiple specific versions of the Docker API can be
supported thanks to the automatic
generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

\textbf{\CRANpkg{AzureContainers}} is an interface to a number of
container-related services in Microsoft's
\href{https://azure.microsoft.com/}{Azure Cloud}
\citep{AzureContainers_2019}. While it is mainly intended for working
with Azure, as a convenience feature it includes lightweight,
cross-platform shells to Docker and Kubernetes (tools \texttt{kubectl}
and \texttt{helm}). These can be used to create and manage arbitrary
Docker images and containers, as well as Kubernetes clusters on any
platform or cloud service.

@@

CARD.com implemented a web application for the optimization of the
acquisition flow and the real-time analysis of debit card transactions.
The software used \CRANpkg{Rserve} and rApache and was deployed in
Docker containers. The R session behind \CRANpkg{Rserve} acted as a
read-only in-memory database, which was extremely fast and scalabale,
for the many concurrent rApache processes responding to the live-scoring
requests of various divisions of the company. Similarly dockerized R
scripts were responsible for the ETL processes and even the
client-facing email, text message and push notification alerts sent in
real-time based on card transactions. The related Docker images were
made available at \url{https://github.com/cardcorp/card-rocker}. The
images extend \texttt{rocker/r-base} and additionally entailed an SSH
client and a workaround for being able to mount SSH keys from the host,
Pandoc, the Amazon Web Services (AWS) SDK, and Java, which is required
by the AWS SDK. The AWS SDK allowed to run R consumers reading from
real-time data processing streams of
\href{https://aws.amazon.com/kinesis/}{AWS Kinesis}
\footnote{See useR!2017 talk \href{https://static.sched.com/hosted\_files/user2017/2f/AWR Kinesis at useR 2017.pdf}{"Stream processing with R in AWS"}.}.
The applications were deployed on Amazon Elastic Container Service
(\href{https://aws.amazon.com/ecs/}{ECS}). The main learnings from using
R in Docker was the imporance of not only pinning the R package versions
via MRAN, but also moving away from Debian testing to a distribution
with long-term support. For the use case at hand, this switch served the
priority to have more control over upstream updates, and to minimize the
risk of breaking the automated builds of the Docker images and
production jobs.

@@

\hypertarget{deployment-and-continuous-delivery}{%
\subsection{Deployment and continuous
delivery}\label{deployment-and-continuous-delivery}}

\label{deployment}

The cloud is the natural environment of containers, and subsequently
containers are the go-to mechanism to deploy R server applications. More
and more continuous integration (CI) and continuous delivery (CD)
services also use containers, opening up new options for use. The
controlled nature of containers, i.e., the possibility to abstract
internal software environment from a minimal dependency outside of the
container is crucial, for example to match test or build environments
with production environments or transfer runnable entities to
as-a-service infrastructures.

First, different packages use containers for the \textbf{deployment of R
and \CRANpkg{Shiny} apps}. \CRANpkg{Shiny} is a popular package for
creating interactive online dashboards with R and it enables users with
very diverse backgrounds to create stable and user friendly web
applications. For example, \emph{ShinyProxy}
(\url{https://www.shinyproxy.io/}) is an open source tool to deploy
Shiny apps in an enterprise context. They feature single sign-on, but
also in scientific use cases
\citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to
achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate
requirements and may itself run in a container for stability and
availability. The tool is build on ContainerProxy
(\url{https://www.containerproxy.io/}), which provides similar features
for executign long-running R jobs or interactive R sessions. The started
containers can run on a regular Docker host but also in clusters.
Another example is the package \CRANpkg{golem}
(\url{https://github.com/ThinkR-open/golem}), which makes an heavy use
of \texttt{dockerfiler} when it comes to creating the
\texttt{Dockerfile} for building production-grade Shiny applications and
deploying them. \CRANpkg{googleComputeEngineR}
(\url{https://cloudyr.github.io/googleComputeEngineR/}) enables quick
deployments of key R services, such as RStudio and Shiny, onto cloud
virtual machines (VMs) with Google Cloud Compute Engine
\citep{googleComputeEngineR_2019}. The package utilises
\texttt{Dockerfile}s to move the labour of setting up those services
from the user to a premade Docker image, which is configured and run in
the cloud VM. For example, by specifying the template
\texttt{template="rstudio"} in functions
\texttt{gce\_vm\_template()/gce\_vm()} an up to date RStudio Service
image is launched for development work, while specifying
\texttt{template="rstudio-gpu"} will launch an RStudio Server image with
a GPU attached, etc.

@@

Third, containers can greatly facilitate the \textbf{deployment to
existing infrastructures}. \emph{Kubernetes}
(\url{https://kubernetes.io/}) is a container-orchestration system for
managing container-based application deployment and scaling. A
\emph{cluster} of containers, orchestrated as a single deployment, e.g.,
with Kubernetes, can mitigate limitations on request volumes or a
container occupied with computationally intensive task. A cluster
features load-balancing, autoscaling of containers across numerous
servers (in the cloud or on premise), and restarting failed ones. It may
be your organisation has a Kubernetes cluster already for other
applications, or you may use one of the different provides for managed
Kubernetes clusters. Docker containers are used within Kubernetes
clusters to hold native code, for which Kubernetes creates a framework
around network connections and scaling of resources up and down. R
applications, big parallel tasks, or scheduled batch jobs can be
deployed in a scalable way using Kubernetes, and the deployment can even
be triggered by changes to code repositories (i.e., CD), see blog post
``R on Kubernetes'' \citep{edmondson_r_2018}. The package
\pkg{googleKubernetesR}
(\url{https://github.com/RhysJackson/googleKubernetesR}) is a proof of
concept for wrapping the Google Kubernetes Engine API, Google's hosted
Kubernetes solution, in an easy to use R package. The package
\CRANpkg{analogsea} provides a solution to programmatically create and
destroy cloud VMs on the \href{https://www.digitalocean.com/}{Digital
Ocean} platform \citep{analogsea_2019}. It also includes R wrapper
functions to install Docker in such a VM, manage images, and control
containers straight from R functions. These functions are translated to
Docker CLI commands and transferred transparently to the respective
remove machine using SSH. \CRANpkg{AzureContainers} is an umbrella
package providing interfaces three commercial services of Microsoft's
Azure Cloud, namely
\href{https://azure.microsoft.com/en-us/services/container-instances/}{Container
Instances} for running individual containers,
\href{https://azure.microsoft.com/en-us/services/container-registry/}{Container
Registry} for private image distribution, and
\href{https://azure.microsoft.com/en-us/services/kubernetes-service/}{Kubernetes
Service} for orchestrated deployments. While a package like
\CRANpkg{plumber} provides the infrastructure for turning an R workflow
a service, for production purposes it is usually necessary to take into
account scalability, reliability and ease of management. AzureContainers
provides an R-based interface to these features and thereby simplifies
complex infrastructure management to a number of R function calls, given
an Azure account with sufficient
credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.

The prevelance of Docker in industry naturally leads to usage of
containers with R in such settings as well, as customers already manage
platforms in Docker containers. These products often entail a high
amount of open source software in combination with proprietary layers
adding the relevant commercializable features. One such example is
RStudio's data science platform
\href{https://rstudio.com/products/team/}{RStudio Team}. It allows teams
of data scientists and their respective IT/DevOps groups to develop and
deploy code in R and Python around the RStudio Open Source Server inside
of Docker images, without requiring users to learn new tools or directly
interact with containers. The best practices for
\href{https://support.rstudio.com/hc/en-us/articles/360021594513-Running-RStudio-with-Docker-containers}{running
RStudio with Docker containers} as well as
\href{https://github.com/rstudio/rstudio-docker-products}{Docker images}
for RStudio's commerical products are publicly available.

@@

The \href{https://mybinder.readthedocs.io/en/latest/}{\textbf{Binder}}
project, maintained by the team behind Jupyter, makes it possible for
users to \textbf{create and share computing environments} with others
\citep{jupyter_binder_2018}. A \emph{BinderHub} allows anyone with
access to a web browser and an internet connection to launch a temporary
instance of these custom environments and execute any workflows
contained within. From a reproducibility standpoint, Binder makes it
exceedingly easy to compile a paper, visualize data, and run small
examples from papers or tutorials without the need for any local
installation. To set up Binder for a project, a user typically starts at
an instance of a BinderHub and passes the location of a repository with
a workspace, e.g., a hosted Git repository, or a data repository like
Zenodo. Binder's core internal tool is \texttt{repo2docker}. It
deterministically builds a Docker image by parsing the contents of a
repository, e.g., project dependency configurations or simple
configuration
files\footnote{See supported file types at \href{https://repo2docker.readthedocs.io/en/latest/config\_files.html}{https://repo2docker.readthedocs.io/en/latest/config\_files.html}.}.
In the most powerful case, \texttt{repo2docker} builds a given
\texttt{Dockerfile}. While this approach works well for most run of the
mill Python projects, it is not so seamless for R projects. This is
partly because \texttt{repo2docker} does not support arbitrary base
images due to the complex auto-generation of the \texttt{Dockerfile}
instructions. Two approaches make using Binder easier. for R users.
First, \textbf{\pkg{holepunch}}
(\url{https://github.com/karthik/holepunch}) is an R package that was
designed to make sharing work environments accessible to novice R users
based on Binder. For any R projects that use the Tidyverse suite
\citep{wickham_welcome_2019}, the time and resources required to build
all dependencies from source can often time out before completion,
making it frustrating for the average R user. \pkg{holepunch} removes
some of these limitations by leveraging Rocker images that contain the
Tidyverse along special Jupyter dependencies, and only installs
additional packages from CRAN and Bioconductor that are not already part
of these images. It short cicuits the configuration file parsing in
\texttt{repo2docker} and starts with the Binder/Tidyverse base images,
which eliminates a large part of the build time and in most cases
results in a binder instance launching within a minute. \pkg{holepunch}
as a side effect also creates a \texttt{DESCRIPTION} file which then
turns any project into a research compendium
\citep{marwick_packaging_2018}. The \texttt{Dockerfile} included with
the project can also be used to launch a RStudio server locally, i.e.,
independent of Binder, which is especially useful when more or special
computational resources can be provided there. The local image usage
reduces the number of seperately managed environments and thereby
reduces work and increases portability and reproducibility. Second, the
\textbf{Whole~Tale} project (\url{https://wholetale.org}) combines the
strengths of the Rocker Project's curated Docker images with
\texttt{repo2docker}. Whole~Tale is a National Science Foundation (NSF)
funded project developing a scalable, open-source, multi-user platform
for reproducible research \citep{brinckman2019, chard2019a}. A central
goal of the platform is to enable researchers to easily create and
publish executable research objects\textbackslash{}footnote\{In
Whole\textasciitilde{}Tale a \emph{tale} is a research object that
contains metadata, data (by copy or reference), code, narrative,
documentation, provenance, and information about the computational
environment to support computational reproducibility\} associated with
published research \citep{chard2019b}. Using Whole~Tale, researchers can
create and publish Rocker-based reproducible research objects to a
growing number of repositories including DataONE member nodes, Zenodo
and soon Dataverse. Additionally, Whole~Tale supports automatic data
citation and is working on capabilities for image preservation and
provenance capture to improve the transparency of published
computational research artifacts \citep{mecum2018, mcphillips2019}. For
R users, Whole~Tale extends the Jupyter Project's \texttt{repo2docker}
component to simplify the customization of R-based environments for
researchers with limited experience with either Docker or Git. Multiple
options have been discussed to allow users to change the base image used
in \texttt{repo2docker} from the default Ubuntu LTS (long-term support)
required to support the Rocker Project images. Whole~Tale implemented a
custom
\texttt{RockerBuildPack}\footnote{See \href{https://github.com/whole-tale/repo2docker\_wholetale}{https://github.com/whole-tale/repo2docker\_wholetale}}
to support customization of the \texttt{rocker/geospatial} image through
\texttt{repo2docker}
composability\footnote{Composability refers to the ability to combine multiple package managers -- such as R, `pip`, and `conda`}.
This works in part because Rocker images are based on a Debian
distribution, so the instructions created by \texttt{repo2docker} for
Ubuntu work because of compatible shell and package manager.

@@

The portability of containerized environments becomes particularly
useful for improving expensive processing of data or shipping complex
processing pipelines. First, it is possible to \textbf{offload complex
processing to a server} or clouds, and also parallel executing of
processes for speeding up or serving of many users.
\textbf{\CRANpkg{batchtools}} provides a parallel implementation of Map
for various schedulers \citep{Lang2017batchtools}. The package can
\href{https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html}{schedule
jobs with Docker Swarm} \textbf{\CRANpkg{googleComputeEngineR}} has the
function \texttt{gce\_vm\_cluster()} to create clusters of 2 or more
virtual machines, running multi-CPU architectures. Instead of running a
local R script with the local CPU and RAM restrictions, the same code
can be processed on all CPU threads of the cluster of machines in the
cloud, all running a Docker container with the same R environments.
\CRANpkg{googleComputeEngineR} integrates with the R parralisation
package \CRANpkg{future} to enable this with only a few lines or R
code\footnote{\href{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}}.
\href{https://cloud.run}{Google Cloud Run} is a CaaS (Containers as a
Service) platform. Users can launch containers using any Docker image
without worrying about underlying infrastructure in a so called
serverless configuration. The service takes care of network ingress,
scaling machines up and down, authentication and authorisation---all
features which are non-trivial for a developer to create on their own.
This can be used to scale up R code to millions of instances if need be
with little or no changes to existing code, as demonstrated by the proof
of concept
\texttt{cloudRunR}\footnote{\href{https://github.com/MarkEdmondson1234/cloudRunR}{https://github.com/MarkEdmondson1234/cloudRunR}},
which uses Cloud Run to create a scalable R-based API using
\CRANpkg{plumber}. \href{https://cloud.google.com/cloud-build/}{Google
Cloud Build} and the Google Container Registry are a continuous
integration service respectively image registry that offload building of
images to the cloud, while serving the needs of commercial environments
such as private Docker images or image stacks. As Googl Cloud Build
itself can run any container, the package \pkg{googleCloudRunner}
demonstrates how R can be used as the control language for one time or
batch processing jobs and scheduling of
jobs\footnote{\href{https://code.markedmondson.me/googleCloudRunner/articles/cloudbuild.html}{https://code.markedmondson.me/googleCloudRunner/articles/cloudbuild.html}}.
\label{drake} \textbf{\CRANpkg{drake}} is a workflow manager for data
science projects \citep{landau_drake_2019}. It features implicit
parallel computing and automated detection of the parts of the work that
actually needs to be reexecuted. \CRANpkg{drake} has been demonstrated
to run inside containers for high
reproducibility\footnote{See for example \href{https://github.com/joelnitta/pleurosoriopsis}{https://github.com/joelnitta/pleurosoriopsis} or \href{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}, the latter even running in a CI, see~\nameref{cicd}.},
but also how to use \CRANpkg{future} package's function
\texttt{makeClusterPSOCK()} to send parts of the workflow to a Docker
image for
execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}}
\citep[see package's function documentation;~][]{future_2020}. In the
latter case, the container control code must be written by the user and
the \CRANpkg{future} package ensures the host and worker can connect for
communicating over socket connections. \textbf{RStudio Server Pro}
includes a functionality called
\href{https://solutions.rstudio.com/launcher/overview/}{Launcher} (since
version~1.2, released in 2019). It gives users the ability to spawn R
sessions and background/batch jobs in a scalable way on external
clusters, e.g.,
\href{https://support.rstudio.com/hc/en-us/articles/360019253393-Using-Docker-images-with-RStudio-Server-Pro-Launcher-and-Kubernetes}{Kubernetes
based on Docker images} or \href{https://slurm.schedmd.com/}{Slurm}
clusters, and optionally, with Singularity containers. A benefit of the
proprietary Launcher software is the ability for R and Python users to
leverage containerisation's advantages in RStudio without writing
